---
title: "Gov 50 Final Project"
author: "Ella Dotzler"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---
https://github.com/fivethirtyeight/data.git

https://github.com/fivethirtyeight/data/blob/cda0c0d9054fc6d020a42f064e87f11c8df85508/hate-crimes/hate_crimes.csv

Introduction:

In modern society, there is a clear baseline intersection of educational output and economic conditions. This will be further analyzed throughout this final project, where the frequency of High School diplomas obtained will be measured in relation to the median household income by area. This research seeks to analyze how correlated the two variables are. The hypothesis during this research is that, as frequency of high school diplomas increases, that the median household income will also increase. Because of this positve relationship, it will likely be the case that regional and community-oriented economic dispar

Data Analysis:





.
```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)

data <- read_csv("hate_crimes_final.csv")
new_names <- as.character(data[1,])
data <- data[-1,]
data <- setNames(data, new_names)

data$median_household_income <- as.numeric(data$median_household_income)
data$share_population_with_high_school_degree <- as.numeric(data$share_population_with_high_school_degree)

histogram <- ggplot(data, aes (x=share_population_with_high_school_degree)) +
  geom_histogram(fill='lightpink',color="grey", bins=30)+
  labs(title="Distribution of High School Degree Share",
       x="Share of Population with High School Degree",
       y="Frequency")
histogram
```


#Results:

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)

data <- read_csv("hate_crimes_final.csv")
new_names <- as.character(data[1,])
data <- data[-1,]
data <- setNames(data, new_names)

data$median_household_income <- as.numeric(data$median_household_income)
data$share_population_with_high_school_degree <- as.numeric(data$share_population_with_high_school_degree)

scatterplot <- data |>
  ggplot(aes(x=median_household_income, y=share_population_with_high_school_degree))+
  geom_point(color='pink', size =4)+
  geom_smooth(method="lm", se=FALSE, color='grey')+
  labs(title="Relationship between Median Household Income and High School Degree Share",x="Median Household Income", y="Share of Population with High School Degree")+
  theme_minimal ()+
  theme(
    plot.title=element_text(size=10, color='black'),
    axis.title=element_text(size=10, color='black'),
    panel.grid = element_blank()
 
  )

scatterplot
```



```{r}
modelsummary <- lm(share_population_with_high_school_degree ~ median_household_income, data = data)

modelsummary
```
#Conclusion
