---
title: "Gov 50 Final Project"
author: "Ella Dotzler"
description: "What is the relationship because educational availability and economic status?"
output:
  distill::distill_article:
    self_contained: false
---

Introduction:

In modern society, there is a clear baseline intersection of educational output and economic conditions. This will be further analyzed throughout this final project, where the frequency of High School diplomas obtained will be measured in relation to the median household income by area. This research seeks to analyze how correlated the two variables are. The hypothesis during this research is that, as frequency of high school diplomas increases, that the median household income will also increase. Because of this positive relationship, it will likely be the case that regional and community-oriented economic disparities will be intertwined with the availability of educational opportunities. This further proves the claim that educational opportunity and availability has a drastic effect on the economic success of both the individual and the community. 

It is also true, though, that the relationship between educational opportunity and economic status is reversible. For example, if one does not have the economic stability, they will fail to fulfill their personal educational needs and that of their offspring—ultimately contributing to a painful cycle. This research aims to contribute to the broader and more meaningful discussion of the connection between economic factors and educational attainment.  

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)

data <- read_csv("hate_crimes_final.csv")
new_names <- as.character(data[1,])
data <- data[-1,]
data <- setNames(data, new_names)

data$median_household_income <- as.numeric(data$median_household_income)
data$share_population_with_high_school_degree <- as.numeric(data$share_population_with_high_school_degree)

histogram <- ggplot(data, aes (x=share_population_with_high_school_degree)) +
  geom_histogram(fill='lightpink',color="grey", bins=30)+
  labs(title="Distribution of High School Degree Share",
       x="Share of Population with High School Degree",
       y="Frequency")
histogram
```

Data Analysis:

This study relies on data sourced from 'fivethirtyeight', a prominent research facility in the United States. This data-set includes information on median household income, the share of the population with a high school degree, and numerous other related factors. Median household income and the share of the population with a high school degree will be the primary focus of the project. This analysis covers block groups following the 2016 presidential election and was initially focused on analyzing post-election hate crimes. However, the data on educational attainment and economic status is equally compelling. 

The outcome variable is the share of the population with a high school degree, and the explanatory variable is the median household income. Other potential variables, such as demographic factors, political affiliation, and employment status could all be large contributers and therefore should not be discounted. However, because the purpose of this research design is to compare median household income to high school degree attainment, they will be overlooked throughout the course of this project.

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)

data <- read_csv("hate_crimes_final.csv")
new_names <- as.character(data[1,])
data <- data[-1,]
data <- setNames(data, new_names)

data$median_household_income <- as.numeric(data$median_household_income)
data$share_population_with_high_school_degree <- as.numeric(data$share_population_with_high_school_degree)

scatterplot <- data |>
  ggplot(aes(x=median_household_income, y=share_population_with_high_school_degree))+
  geom_point(color='pink', size =4)+
  geom_smooth(method="lm", se=FALSE, color='grey')+
  labs(title="Relationship between Median Household Income and High School Degree Share",x="Median Household Income", y="Share of Population with High School Degree")+
  theme_minimal ()+
  theme(
    plot.title=element_text(size=10, color='black'),
    axis.title=element_text(size=10, color='black'),
    panel.grid = element_blank()
  )
scatterplot
```
Results:

The scatter plot created above shows a clearly positive relationship between median household income and the share of the population with a high school degree. This visual depiction of the data suggests a positive relationship between income and educational attainment, meaning that as income increases, high school degree attainment will as well. What is most interesting about this data set though, is that it is not explicitly clear as to which direction the positive relationship flows. For example, it could also be the case that, as educational attainment increases, median household income will in response. That is what makes this data set particularly worth studying—the variables are so clearly interwined that the correlation and causation becomes a bit complicated. 


```{r}

library(tidyverse)
library(broom)
library(kableExtra)

modelsummary <- lm(share_population_with_high_school_degree ~ median_household_income, data = data)

modelsummary

tidy_results <- tidy(modelsummary) 

regression_table <- tidy_results |>
  kable(format="html") 

regression_table

```

Conclusion:

Data Information:

https://github.com/fivethirtyeight/data.git

https://github.com/fivethirtyeight/data/blob/cda0c0d9054fc6d020a42f064e87f11c8df85508/hate-crimes/hate_crimes.csv
